---
title: "Data preparation"
output: html_document
date: "2024-10-31"
bibliography: references.bib
author: "Lucan Bovens"
---

```{=html}
<style>
body {
text-align: justify}
</style>
```
  
```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
----

# Data preparation
  
----
  
## Prep

Start with a clean workspace to prepare all the data.

```{r}
rm(list=ls())
```

## Functions

Activating the user defined functions. Thanks to Jochem Tolsma for providing these during the course.

```{r}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    totalname <- paste(location, file, ".rda", sep = "")
    save(x, file = totalname)
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}
```

```{r}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id #openalex IDs
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]
    
    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}
```

## Loading data

Activating the packages needed.

```{r}
packages <- c("tidyverse", "scholar", "openalexR", "rvest", "jsonlite", "RSiena", "RSelenium", "rvest", "lavaan", "xml2", "kableExtra", "devtools", "igraph", "network", "sna", "ggraph", "visNetwork", "threejs", "networkD3", "RsienaTwoStep")

fpackage.check(packages)
```

Enabling the connection to OpenAlex API.

```{r}
# make yourself known to openAlex
options(openalexR.mailto = "lucan.bovens@home.nl")
```

Loading the original datafile.

```{r}
scholars <- fload("./data/processed/scholars_20240924.rda")
```

Making sure to save the three Sociology departments. I chose Radboud University (RU), Utrecht University (UU) and the University of Groningen. There are two waves: 1. 2015-2018; 2. 2019-2023. The type of ties is first, meaning the first author is assumed to be the one to initiate ties.

```{r}

#Save the output of your function, sociology departments
socdata <- fcolnet(data = scholars, 
                university = c("RU", "UU", "RUG"), 
                discipline = c("sociology"), 
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first")) #First instead of all to check direction of ties

```

Finding the ego characteristics (df_ego).

```{r}

#Let us find ego characteristics. 
#first fish out the data
df <- socdata$data

#same complicated structure as 'scholars' thus first make a dataframe from the list in which all info was saved. 
df_ego <- do.call(rbind.data.frame, df$demographics)

#DO NOT MESS UP THE ORDER! THUS IF YOU JOIN THIS DATA WITH YOUR OWN DATA CHECK THAT ORDER REMAINED THE SAME!! 

```

```{r}
View(df_ego)
```

```{r}
demographics <- do.call(rbind.data.frame, scholars$demographics)

universities_1_2022 <- unique(demographics$Universiteit1.22)
universities_2_2022 <- unique(demographics$Universiteit2.22)
universities_1_2024 <- unique(demographics$Universiteit1.24)
universities_2_2024 <- unique(demographics$Universiteit2.24)

```

## Adding variables to dataset
### H-index

The following syntax adds the H-index, as well as some other metrics such as the i10-index, the number of works each authors has and the total times an author has been cited by others. The (adjusted) H-index is the main variable for this study.

```{r}

# Add h-index and i10-index to df_ego dataframe
df_ego <- df_ego %>%
  mutate(h_index = NA_real_, i10_index = NA_real_, works_count = NA_real_, cited_by_count = NA_real_)

# Get H-index and i10-index, using openalex API (see link)
# Additionally get works count and cited_by count for possible future use
get_openalex_metrics <- function(openalex_id) {
  url <- paste0("https://api.openalex.org/authors/", openalex_id)
  author_data <- tryCatch({
    jsonlite::fromJSON(url)
  }, error = function(e) {
    message(paste("Error fetching data for OpenAlex ID:", openalex_id)) #After googling for a bit, found this to know if something is running forever or whether errors show up
    return(NULL)
  })
  
  if (!is.null(author_data)) {
    h_index <- author_data$summary_stats$h_index
    i10_index <- author_data$summary_stats$i10_index
    works_count <- author_data$works_count
    cited_by_count <- author_data$cited_by_count
    return(list(h_index = h_index, i10_index = i10_index, works_count = works_count, cited_by_count = cited_by_count))
  } else {
    return(list(h_index = NA, i10_index = NA, works_count = NA, cited_by_count = NA))
  }
}

# Loooop (may take a while)
for (i in 1:nrow(df_ego)) {
  openalex_id <- df_ego$au_id[i]  
  metrics <- get_openalex_metrics(openalex_id)
  df_ego$h_index[i] <- metrics$h_index
  df_ego$i10_index[i] <- metrics$i10_index
  df_ego$works_count[i] <- metrics$works_count
  df_ego$cited_by_count[i] <- metrics$cited_by_count
}

View(df_ego)

```

### Gender

Thanks to the gender-scraper developed by Niels Vullings, I was able to add gender as a control variable.

```{r}
# Credits to Niels Vullings

# Extract first names from df_ego
df_ego <- df_ego %>%
  mutate(first_name = sapply(strsplit(Naam, " "), `[`, 1))

# Connection errors may occur, if so, restart

gender_scraper.NV <- function(names = "names element", web_page = "https://nvb.meertens.knaw.nl/naam/is/") {
  
  names$first_name <- sapply(strsplit(names$Naam, " "), `[`, 1)  # Extract first names from the ego characteristics dataset
    names$male <- NA
  names$male <- NA
  names$female <- NA
  
  for(i in 1:nrow(names)) {
    
    web_page <- read_html(paste0("https://nvb.meertens.knaw.nl/naam/is/", names$first_name[i]))
    
    table <- web_page %>% 
      rvest::html_elements("body") %>% 
      rvest::html_elements("table") %>% 
      rvest::html_table()
    
    if(length(table) == 0) {
      names$male[i] <- NA
      names$female[i] <- NA
      
    } else {
      names$male[i] <- as.numeric(ifelse(table[[1]][[2,3]] == "--", 0, table[[1]][[2,3]]))  
      names$female[i] <- as.numeric(ifelse(table[[1]][[6,3]] == "--", 0, table[[1]][[6,3]]))
    }
  }  # End for-loop
  
  names <- names %>% mutate(perc_female = case_when(is.na(female) & is.na(male) ~ NA,
                                                    is.na(female) ~ 0,
                                                    is.na(male) ~ 1,
                                                    TRUE ~ round((female/(male + female)), 2))) %>% 
    select(-male, -female)  # Remove intermediate columns
  
  return(names)
}

# Run gender scraper -> df_ego
df_ego <- gender_scraper.NV(names = df_ego, web_page = "https://nvb.meertens.knaw.nl/naam/is/")

View(df_ego)

# Set perc_female to 0 for 'Weverthon'
df_ego <- df_ego %>%
  mutate(perc_female = ifelse(first_name == "Weverthon", 0, perc_female))

# Set perc_female to 1 for 'Xingna'
df_ego <- df_ego %>%
  mutate(perc_female = ifelse(first_name == "Weverthon", 1, perc_female))

check <- df_ego %>% count(perc_female)
plot(check$perc_female, check$n)

```

### Gender create dichotomous

I decided to make gender a dichotomous variable to make it clearer.

```{r}
# Dichotomous gender?
df_ego <- df_ego %>%
  mutate(genderdich = case_when(
    perc_female > 0.5 ~ "female",
    perc_female <= 0.5 ~ "male",
    TRUE ~ NA_character_
  ))


table(df_ego$genderdich)

# Remove authors gender is unknown (NA value)
df_ego <- df_ego %>% 
  filter(!is.na(genderdich))

View(df_ego)

table(df_ego$genderdich)
```

### Career age

Adding the career age to serve as a factor when creating the new prestige-index. The below syntax was developed by Koen Lucas, who I thank.

```{r}

df_works <- socdata$data$work

# Handy: unnest df_works to get the correct format
df_works <- tibble(works_id = unlist(lapply(df_works, function(l) l$id)),
                   works_author = unlist(lapply(df_works, function(l) l$author), recursive = FALSE),
                   works_year = unlist(lapply(df_works, function(l) l$publication_year), recursive = FALSE))

#Continue
df_career_age <- df_works %>%
  unnest(works_author) %>%  # Unnest tibble (authors) from the (df_works) tibble
  filter(au_id %in% df_ego$au_id == TRUE) %>%  # authors in df_works should match authors in df_ego
  group_by(au_id) %>%
  mutate(min_year_pub = min(works_year, na.rm = TRUE),  # Initial calculation minimum year
         count_pub = n()) %>%  # Count  number of publications
  mutate(first_year_pub = ifelse(count_pub == 1, min_year_pub,  # if number of publications = 1, score is min_year_pub
                                 {
                                   filter_year = works_year[works_year != min_year_pub]  # Filter  min year
                                   mean_pubyear = mean(filter_year, na.rm = TRUE)  # Calculate mean excluding min score
                                   valid_year = min(filter_year[filter_year >= (mean_pubyear - 25)], na.rm = TRUE)  # Min year ->  range mean-25
                                   valid_year  # if number of publications is not 1, then score minimum year with first year removed within range mean-25
                                 })) %>%
  
  # this line corrects for the warning of returning infinites that for some reason were not filtered through the first ifelse command
  mutate(first_year_pub = ifelse(is.infinite(first_year_pub), min_year_pub, first_year_pub)) %>%
  
  select(au_id, first_year_pub) %>%
  distinct(au_id, first_year_pub, .keep_all = TRUE)

# Merge dataframed based on au_id
df_ego <- df_ego %>%
  left_join(df_career_age, by = "au_id") %>%
  mutate(career_age = 2024 - first_year_pub)  # Calculate career age as the difference from 2024

View(df_ego)

```

### Creating new prestige variable

I decided to create a new prestige variable which takes career age into account as well. I used the logarithm to get to fairer results.

```{r}
# If career age 0, change to 1 (to solve the dividing by 0 problem)
df_ego <- df_ego %>%
  mutate(adjusted_career_age = ifelse(career_age == 0, 1, career_age))
```


```{r}
# Calculate prestige by dividing h-index by the log of career age + 1 (to avoid division by zero problems)
df_ego <- df_ego %>%
  mutate(prestige_index = h_index / log1p(adjusted_career_age))

View(df_ego)
```

## Missings

I excluded a few missings which made no sense and which could not be used. 

```{r}
df_ego <- df_ego %>%
  filter(!(works_count == 0 & !is.na(adjusted_career_age)))
```

## Some final prep

```{r}
final_scholar_ids <- df_ego$au_id

df <- list(
  demographics = df_ego %>% select(au_id, everything()),  
  scholars_oa = df_ego %>% select(au_id, h_index, i10_index), 
  works = df_ego %>% select(au_id, works_count, cited_by_count) 
)

sapply(df, nrow)  # Should show 164 for each component instead of 171 due to excluding missings
```

```{r}
View(df_ego)
```

## Saving the data

Saving the data so I can use it for the descriptives and analyses.

```{r}
# Saving the data
fsave(df_ego, "df_ego_final")
fsave(socdata, "socdata_final")
fsave(df_works, "df_works_final")
```

